On the probability of finding a nonphysical solution through shadowing

### 2020-10-27 main.tex
plot_map update by qiqi
Add Sauer reference.

In view of the plucked tent map, we can question further the
validity of numerical solutions. Can numerical play 
the role of the plucking perturbation to the 
governing equation? That is, can numerical solutions represent 
slight perturbations to the governing equations that nevertheless 
exhibit drastically different statistics from that of the ensemble? Uniformly hyperbolic 
dynamical systems exhibit \emph{linear response} 
(\cite{ruelle}\cite{baladi}) by which small parameter perturbations 
lead to small changes in statistics, which can be expanded as Taylor series around the reference parameter value. But, uniform hyperbolicity is a mathematical idealization, and although some physical systems have been observed to behave as if they were uniformly hyperbolic \cite{gallavotti}, a violation of uniform hyperbolicity is more likely 
\cite{wormell}\cite{luca}. The plucked tent map provides a pathological example of a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).

@@

Consider $s=0.5$. 
One way to see the asymmetry in the probability distribution 
about $x=1$, at $n=0$, is to construct a Markov chain with nodes 00, 01, 10, and 11, that indicate the subintervals $(0,0.5)$, $[0.5,1)$, 
$[1,1.5)$ and $[1.5,2)$, respectively. From Eq. \ref{eqn:plucked}, we can see that the application of $\varphi_{0.5,0}$ yields the following transition matrix:
\[
\begin{bmatrix}
		1/4 & 3/4 & 0 & 0 \\
		0 & 0 & 3/4 & 1/4 \\
		0 & 0 & 3/4 & 1/4 \\
		1/4 & 3/4 & 0 & 0
\end{bmatrix}
.\] 
The stationary probability distribution of this Markov chain is $[1/8, 3/8, 9/8, 3/8]^T$, consistent with the left hand side of Figure \ref{fig:pluckedTentMapDistribution}. Clearly, the left subinterval, $[0,1)$ is three times less likely to be visited by a long trajectory of $\varphi_{0.5,0}$, compared to $[1,2)$. In fact, the probability of visiting $[0,1)$ remains at 0.25 for all values of $n$, but for $n >> 0$, fractal distributions are produced within smaller subsets of the two subintervals. 
\end{document}

>>>>

In view of the plucked tent map, we can question further the
validity of numerical solutions. Can numerical play 
the role of the plucking perturbation to the 
governing equation? That is, can numerical solutions represent 
slight perturbations to the governing equations that nevertheless 
exhibit drastically different statistics from that of the ensemble? Uniformly hyperbolic 
dynamical systems exhibit \emph{linear response} 
(\cite{ruelle}\cite{baladi}) by which small parameter perturbations 
lead to small changes in statistics, which can be expanded as Taylor series around the reference parameter value. But, uniform hyperbolicity is a mathematical idealization, and although some physical systems have been observed to behave as if they were uniformly hyperbolic \cite{gallavotti}, a violation of uniform hyperbolicity is more likely 
\cite{wormell}\cite{luca}. Sauer \cite{sauer} has shown 
examples in which numerical error due to computation in double-precision floating point arithmetic cause significant change in the stationary probability distribution on the attractor. The plucked tent map adds to this list of examples by exhibiting a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).

@@

Consider $s=0.5$. 
One way to see the asymmetry in the probability distribution 
about $x=1$, at $n=0$, is to construct a Markov chain with nodes 00, 01, 10, and 11, that indicate the subintervals $(0,0.5)$, $[0.5,1)$, 
$[1,1.5)$ and $[1.5,2)$, respectively. From Eq. \ref{eqn:plucked}, we can see that the application of $\varphi_{0.5,0}$ yields the following transition matrix:
\[
\begin{bmatrix}
		1/4 & 3/4 & 0 & 0 \\
		0 & 0 & 3/4 & 1/4 \\
		0 & 0 & 3/4 & 1/4 \\
		1/4 & 3/4 & 0 & 0
\end{bmatrix}
.\] 
The stationary probability distribution of this Markov chain is $[1/8, 3/8, 9/8, 3/8]^T$, consistent with the left hand side of Figure \ref{fig:pluckedTentMapDistribution}. Clearly, the left subinterval, $[0,1)$ is three times less likely to be visited by a long trajectory of $\varphi_{0.5,0}$, compared to $[1,2)$. In fact, the probability of visiting $[0,1)$ remains at 0.25 for all values of $n$, but for $n >> 0$, fractal distributions are produced within smaller subsets of the two subintervals. The scripts that generate the plucked tent map and its 
stationary probability distributions can be found under \verb+tent_sens_stat+. 
\end{document}
